###Packages
```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(nflfastR)
library(readr)
library(scales)
library(tidyverse)
library(tidyr)
library(zoo)
```

###Grid approximation
```{r}
###Pick season and load data
seasons <- 2015:2020
pbp <- purrr::map_df(seasons, function(x) {
  readRDS(
    url(
      glue::glue("https://raw.githubusercontent.com/guga31bb/nflfastR-data/master/data/play_by_play_{x}.rds")
    )
  )
})

####Function to use grid approximation for player of choice and mean of choice (I am using 2020 league average CP)
Grid_Approximation_Function <- function(QB_name){

###Get QB attempts
QB_Attempts <- pbp %>%
  filter(passer_player_name == QB_name) %>%
  filter(sack == 0) %>%
  select(desc, complete_pass, incomplete_pass) %>%
  nrow()

###Get QB completions
QB_Completions <- pbp %>%
  filter(passer_player_name == QB_name) %>%
  filter(sack == 0) %>%
  select(desc, complete_pass, incomplete_pass) %>%
  summarise(complete = sum(complete_pass)) 

###Create a grid of possible values for completion percentage
p_grid <- seq(from = 0, to = 1, length.out = 200)

###create a flat prior to start
prior <- rep(1,200)

###Compute likelihood
likelihood <- dbinom(QB_Completions$complete[[1]], size = QB_Attempts, prob = p_grid)

###Compute product
unstd.posterior <- likelihood * prior

###standardize the posterior to sum to 1
posterior <- unstd.posterior / sum(unstd.posterior)

###League average CP, I chose 2020 
league_mean_CP <- pbp %>%
  filter(is.na(epa) == FALSE) %>%
  filter(play_type == "pass") %>%
  filter(play_type == "pass" , season == 2020) %>%
  select(desc, complete_pass, incomplete_pass) %>%
  filter(is.na(complete_pass)==FALSE) %>%
  summarise(mean = mean(complete_pass))

###Convert grid into data frame for usage in GGPLOT
p_grid_graphing <- p_grid %>%
  as.data.frame() %>%
  rename("p_grid" = ".")

###Convert posterior into data frame for usage in GGPLOT
posterior_graphing <- posterior %>%
  as.data.frame() %>%
  rename("posterior" = ".")

###Viz
return(
cbind(p_grid_graphing, posterior_graphing)  %>%
  ggplot(aes(x=p_grid, y=posterior)) +
  theme_light()+
  theme(plot.title = element_text(color="black", size=8, face="bold"))+
  coord_cartesian(clip = "off") +
  theme(plot.title = element_text(size = 10, face = "bold"),
  plot.subtitle = element_text(size = 8))+
  theme(plot.background = element_rect(fill = "gray97"))+
  theme(panel.background = element_rect(fill = "gray97"))+
  geom_line() +
  geom_vline(xintercept = league_mean_CP$mean[[1]], linetype = "dashed", color = "dark red") +
  coord_cartesian(clip = "off") +
  theme(plot.title = element_text(size = 10, face = "bold"),
  plot.subtitle = element_text(size = 8))+
  labs(title = paste0(QB_name, " Bayesian Grid Approximation of Completion Percentage"),
       caption = "Plot: PatriotsStatsR, Data: NFLFastR")+
    ylab("Posterior Probability")+
  xlab("Completion Percentage") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_x_continuous(labels = percent_format(accuracy = 1), breaks = seq(0,1,.05))

)

}

Grid_Approximation_Function("J.Stidham") %>%
  ggsave(filename = "QB_Bayes.png", 
       dpi = 300, type = "cairo", width = 10, height = 7, units = "in")

```

###Loop to compare quarterbacks
```{r}
####Function to use grid approximation for player of choice and mean of choice (I am using 2020 league average CP)
Grid_Approximation_Compare_Function <- function(QB_name){

###Get QB attempts
QB_Attempts <- pbp %>%
  filter(passer_player_name == QB_name) %>%
  filter(sack == 0) %>%
  select(desc, complete_pass, incomplete_pass) %>%
  nrow()

###Get QB completions
QB_Completions <- pbp %>%
  filter(passer_player_name == QB_name) %>%
  filter(sack == 0) %>%
  select(desc, complete_pass, incomplete_pass) %>%
  summarise(complete = sum(complete_pass)) 

###Create a grid of possible values for completion percentage
p_grid <- seq(from = 0, to = 1, length.out = 200)

###create a flat prior to start
prior <- rep(1,200)

###Compute likelihood
likelihood <- dbinom(QB_Completions$complete[[1]], size = QB_Attempts, prob = p_grid)

###Compute product
unstd.posterior <- likelihood * prior

###standardize the posterior to sum to 1
posterior <- unstd.posterior / sum(unstd.posterior)

###Convert grid into data frame for usage in GGPLOT
p_grid_graphing <- p_grid %>%
  as.data.frame() %>%
  rename("p_grid" = ".")

###Convert posterior into data frame for usage in GGPLOT
posterior_graphing <- posterior %>%
  as.data.frame() %>%
  rename("posterior" = ".")

###Bind our data and label our QB
return(
QB_CP_Data  <- cbind(p_grid_graphing, posterior_graphing)
)

}

###Blank data frame for merging
QB_CP_Database <- data.frame()

###List of QBS you are interested in 
QBs_Of_Interest <- list("J.Stidham",
                        "C.Newton",
                        "B.Hoyer")

###Loop to get values
for (x in 1:length(QBs_Of_Interest)) {
  QB_Data <- Grid_Approximation_Compare_Function(QBs_Of_Interest[x]) %>%
    mutate(QB = QBs_Of_Interest[[x]])
  QB_CP_Database <- rbind(QB_CP_Database, QB_Data)
}

###Viz
QB_CP_Database %>%
  ggplot(aes(x=p_grid, y=posterior, group = QB)) +
  theme_light()+
  theme(plot.title = element_text(color="black", size=8, face="bold"))+
  coord_cartesian(clip = "off") +
  theme(plot.title = element_text(size = 10, face = "bold"),
  plot.subtitle = element_text(size = 8))+
  theme(plot.background = element_rect(fill = "gray97"))+
  theme(panel.background = element_rect(fill = "gray97"))+
  geom_line(aes(linetype = QB, color = QB)) +
  geom_vline(xintercept = 0.6124461, linetype = "dashed", color = "dark red") +
  coord_cartesian(clip = "off") +
  theme(plot.title = element_text(size = 10, face = "bold"),
  plot.subtitle = element_text(size = 8))+
  labs(title = paste0( " Bayesian Grid Approximation of Completion Percentage"),
       caption = "Plot: PatriotsStatsR, Data: NFLFastR")+
    ylab("Posterior Probability")+
  xlab("Completion Percentage") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_x_continuous(labels = percent_format(accuracy = 1), breaks = seq(0,1,.05)) +
  ggsave(filename = "QB_All.png", 
       dpi = 300, type = "cairo", width = 10, height = 7, units = "in")
  



