###Packages
```{r}
library(cfbfastR)
library(dplyr)
library(ffscrapr)
library(tidyverse)
library(hrbrthemes)
library(rethinking)
library(viridis)
library(viridisLite)
library(brms)
library(zoo)
library(nflfastR)
library(Lahman)
library(stats4)
library(VGAM)
library(gamlss)
library(broom)
```

###Load in data and clean 
```{r}
#load data and remove NA values
QB_Data <- readRDS("CFB_QB_Data.rds") %>%
  filter(is.na(EPA)    == FALSE) %>%
  filter(is.na(stars)  == FALSE) %>%
  filter(is.na(rating) == FALSE) %>%
  filter(pass_attempt  == 1)

#make stars a factor variable
QB_Data$stars <- as.factor(QB_Data$stars)

#round rating
QB_Data$rating <- round(QB_Data$rating, digits = 2)

QB_Data$start_date <- as.Date(QB_Data$start_date , format = "%m/%d/%Y")
```

###Explore what predicts a QB's career best, stars, composite, rating
```{r}
#Stars
QB_Data %>%
  group_by(stars, name) %>%
  summarize(mean_cp = mean(completion, na.rm = TRUE),
            std_cp  = sd(completion, na.rm = TRUE),
            throws  = n()) %>%
  filter(throws > 50) %>%
  ungroup() %>%
  group_by(stars) %>%
  summarize(mean_cp = mean(mean_cp, na.rm = TRUE),
            std_cp  = sd(std_cp, na.rm = TRUE))

#Composite
QB_Data %>%
  group_by(rating, name) %>%
  summarize(mean_cp = mean(completion, na.rm = TRUE),
            std_cp  = sd(completion, na.rm = TRUE),
            throws  = n()) %>%
  filter(throws > 50) %>%
  ungroup() %>%
  group_by(rating) %>%
  summarize(mean_cp = mean(mean_cp, na.rm = TRUE),
            std_cp  = sd(std_cp, na.rm = TRUE))

#Ranking
QB_Data %>%
  group_by(ranking, name) %>%
  summarize(mean_cp = mean(completion, na.rm = TRUE),
            std_cp  = sd(completion, na.rm = TRUE),
            throws  = n()) %>%
  filter(throws > 50) %>%
  ungroup() %>%
  group_by(ranking) %>%
  summarize(mean_cp = mean(mean_cp, na.rm = TRUE),
            std_cp  = sd(std_cp, na.rm = TRUE))
```

###I chose to use the composite rating because it adds a little more specificity than just a factor level star rating. We can viz it here with a smoothing spline
```{r}
QB_Data %>%
  group_by(rating, name) %>%
  summarize(mean_cp = mean(completion, na.rm = TRUE),
            std_cp  = sd(completion, na.rm = TRUE),
            throws  = n()) %>%
  filter(throws > 50)  %>%
  ggplot(aes(x=rating, y=mean_cp)) +
  geom_point() +
  geom_smooth()
```

###I will use a beta prior. Two things we need to account for in our model. 1.)The relationship between completions and throw attempts. A player who performs poorly will get less chances and a player who performs well will get more chances & 2.) Experience of a QB, if a QB is a 4th or 5th year starter our prior should shift to a higher expectation while a freshman should have a lower expectation
```{r}
##visualize relationship 1. just mentioned with a log scale, as the number of throws increases so generally does the completion percentage. So we need to account for this
QB_Data %>%
  group_by(id, name) %>%
  summarize(throws = n(),
            cp = sum(completion) / sum(pass_attempt)) %>%
  ggplot(aes(x = throws, y = cp)) +
  geom_point() +
  geom_hline(yintercept = 0.5524417) +
  geom_smooth(method = "lm") +
  scale_x_continuous(trans='log2')

##visualize relationship 2 just mentioned, no need for log scale (instead will use second degree polynomial but we can see that freshman should have a different prior than other QBs and so on. A bit of a plateau once a QB reaches year 3. 
QB_Data %>%
  mutate(Years_Since_Entry = ifelse(Season   == class, 0, 
                             ifelse(Season == class+1, 1,
                             ifelse(Season == class+2, 2,
                             ifelse(Season == class+3, 3,
                             ifelse(Season == class+4, 4, 5)))))) %>%
  dplyr::select(season, class, Years_Since_Entry, completion, pass_attempt, rating, id) %>%
  group_by(Years_Since_Entry) %>%
  summarize(cp = mean(completion),
            ex = n()) %>%
  arrange(desc(Years_Since_Entry)) %>%
  ggplot(aes(x=Years_Since_Entry, y = cp)) +
  geom_point() +
  stat_spline(nknots = 6) + #check different spline values, all seem very similar
  stat_spline(nknots = 5) +
  stat_spline(nknots = 4)
```

###Fit model and account for those biases
```{R}
##Data to fit
QB_Grouped <- QB_Data %>%
  mutate(Years_Since_Entry = ifelse(Season   == class, 0, 
                             ifelse(Season == class+1, 1,
                             ifelse(Season == class+2, 2,
                             ifelse(Season == class+3, 3,
                             ifelse(Season == class+4, 4, 5)))))) %>%
  group_by(id, name, rating, Years_Since_Entry) %>%
  summarize(throws      = n(),
            cp          = sum(completion) / sum(pass_attempt),
            completions = sum(completion),
            attempts    = sum(pass_attempt)) %>%
  ungroup()

##fit gam lss
fit <- gamlss(cbind(completions, attempts - completions) ~ 0 + log(attempts) + rating + ns(Years_Since_Entry, df = 4),
              data = QB_Grouped,
              family = BB(mu.link = "identity"))

tidy(fit)

#fit model to data
mu    <- fitted(fit, parameter = "mu")
sigma <- fitted(fit, parameter = "sigma")

career_eb_wCP <- QB_Grouped %>%
  dplyr::select(name, id, completions, attempts, rating, Years_Since_Entry) %>%
  mutate(mu = mu,
         alpha0 = mu / sigma,
         beta0 = (1 - mu) / sigma,
         alpha1 = alpha0 + completions,
         beta1 = beta0 + attempts - completions,
         new_eb = alpha1 / (alpha1 + beta1),
         low = qbeta(.025, alpha1, beta1),
                high = qbeta(.975, alpha1, beta1)) 
```

#Function to plot player of interest
```{r}

QB_Graphing_Function <- function(QB_Name, color_code) {
a <- QB_Data %>%
  mutate(Years_Since_Entry = ifelse(Season   == class, 0, 
                             ifelse(Season == class+1, 1,
                             ifelse(Season == class+2, 2,
                             ifelse(Season == class+3, 3,
                             ifelse(Season == class+4, 4, 5)))))) %>%
  filter(name == QB_Name) %>%
  arrange(start_date, game_play_number) %>%
  within(acc_sum_cp <- cumsum(completion)) %>%
  within(acc_sum_at <- cumsum(pass_attempt))


b <- career_eb_wCP %>%
  arrange(desc(new_eb)) %>%
  filter(  name == QB_Name  ) %>%
  dplyr::select(-new_eb,
         -low,
         -high)

return(merge(a,b, by = c("name", "id", "Years_Since_Entry")) %>%
  mutate(alpha1 = alpha0 + acc_sum_cp,
         beta1 = beta0 + acc_sum_at - acc_sum_cp,
         new_eb = alpha1 / (alpha1 + beta1),
         low = qbeta(.025, alpha1, beta1),
                high = qbeta(.975, alpha1, beta1))%>%
  ggplot(aes(x=acc_sum_at, y=new_eb)) +
  geom_line(color = color_code)+
  geom_ribbon(aes(ymin=low, ymax=high), linetype=2, alpha=0.1, color = "#828A8F") +
    xlab("Throw Attempt") +
  ylab("Completion Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_light()+
  theme(plot.title = element_text(color="black", size=8, face="bold"))+
  theme(plot.title = element_text(size = 10, face = "bold"),
  plot.subtitle = element_text(size = 8))+
  theme(plot.background = element_rect(fill = "gray97"))+
  theme(panel.background = element_rect(fill = "gray97"))+
  labs(title = "Bayesian Updating of Completion Percentage with Credible Intervals: Recruit Rating, 
Number of Dropbacks & Years of Experience as Prior",
       caption = "Plot: PatriotsStatsR, Data: CFBFastR") +
  geom_hline(yintercept = quantile(quantile(QB_Grouped$cp, .5)), linetype = "dashed") +
  geom_hline(yintercept = quantile(quantile(QB_Grouped$cp, .75)), linetype = "dashed") +
  geom_hline(yintercept = quantile(quantile(quantile(QB_Grouped$cp, .9)[1])), linetype = "dashed") +
  annotate("text", x = 525, y = .74, label = "90th Percentile", size = 3) +
  annotate("text", x = 525, y = .63, label = "75th Percentile" , size = 3) +
  annotate("text", x = 525, y = .57, label = "50th Percentile", size = 3))
    ggsave(filename = paste0("Mac_Updating_wCP",".png"), 
       dpi = 300, type = "cairo", width = 10, height = 7, units = "in")
}

#insert QB name and the color of the line desired
QB_Graphing_Function("Kellen Mond", "#9E1B32")
```





